---
title: "GITM"
output: html_document
date: "2025-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Library Imports
```{r}
library(hisse) # Misse
library(phytools)
library(ape)
#library(geiger)
library(ggplot2)
library(dplyr)
library(giscoR)
#library(diversitree)
#library(BAMMtools)
```

# Reading in Tree Data
```{r}
myTree2 <- read.nexus(file="Supplementary File 2.tre")
plotTree(myTree2,fsize = .3)
myTree3 <- read.nexus(file="Supplementary File 3.tre")
plotTree(myTree3,fsize= .3)
myTree4 <- read.nexus(file="Supplementary File 4.tre")
plotTree(myTree4, fsize=0.3)
```

```{r}
# some code sourced from Paige Cherry in this code chunk
# uploaded the data from the 2016 files into R
distribution <- read.csv("Distribution.csv")
bodysize <- read.csv("BodySize.csv")
# uploaded (sara) my research data about islands
island <- read.csv('SpeciesIsland.csv', skip = 1)

# merged distribution and island size
chartdf <- merge(distribution, bodysize, by="Species")
chartdf2 <- merge(chartdf,island, by="Species")

# separated the merged data frame from where the tortoises lived
#mainland <- chartdf[which(chartdf[,3]==0),]
#shared <- chartdf[which(chartdf[,3]==1),]
#islands <- chartdf[which(chartdf[,3]==2),]

# created a vector (list of #s or names) for them
#meanvec <- c(mean(mainland[,4]), mean(islands[,4]), mean(shared[,4]))
#namevec <- c("mainland", "islands", "shared")

# plotted them on a bar chart
# barplot(meanvec, names.arg=namevec, main="mean values of body size")
# 
#stanvec <- c(sd(mainland[,4]), sd(islands[,4]), sd(shared[,4]))
# barCenters <- barplot(height = meanvec, names.arg=namevec,
#                       main = "mean values of body size", xlab = "Group", ylab = "bodysize", 
#                       ylim = c(0,150))
# arrows(barCenters, meanvec-stanvec,
#        barCenters, meanvec+stanvec,angle=90,code=3)

```

writing out the chart for Hubert for website will add the tip rates on later and write out again
```{r}
write.csv(chartdf2,'all_data.csv')
```


# BOXPLOT: original data distribution
Mainland, Shared, Island
plotting the box plots of weight by location
```{r}
# create list for area
area = c('mainland','shared','island')
# set location in the chart to area based on index position
chartdf2$location = area[(chartdf2$Islands..multistate..x +1)]
# set as a factor
chartdf2$location = as.factor(chartdf2$location)
# plotting box plots for body length by locations
ggplot(chartdf2, aes(x=location, y=Carapace.length..cm.)) + 
  geom_boxplot(fill = '#42a5f5')+
  labs(y= "Body Length (cm)", x = "Groups", title ='Body Size Across Groups' )
```
# Violin plot: original data
same plot as above visualized differently 
did not end up getting used on the website or poster
```{r}
#plot violin plots for the body length by location
ggplot(chartdf2, aes(x=location, y=Carapace.length..cm.)) + 
  geom_violin(fill = '#66bb6a')+
  labs(y= "Body Length (units)", x = "Groups", title ='Body Size Across Groups' )
```
# BOX: Galapagos vs. Other
sorted out data by Galapagos and other region to plot box plot
this gives support to the fact that the Galapagos tortoises are large and give background 
context to the island plot by showing that the Galapagos tortoises plotted are larger 
than most other species
```{r}
galapagos = c('Other','Galapagos')
# set location in the chart to area based on index position
chartdf2$galapagos = galapagos[chartdf2$Galápagos..binary..y +1]
chartdf2$galapagos = as.factor(chartdf2$galapagos)
# plotting box plots for body length by locations
ggplot(chartdf2, aes(x=galapagos, y=Carapace.length..cm.)) + 
  geom_boxplot(fill = '#66bb6a')+
  labs(y= "Body Length (cm)", x = "Groups", title ='Body Size Across Groups' )
```

# MAP
```{r}
## https://r-graph-gallery.com/330-bubble-map-with-ggplot2.html
#island locations: https://datazone.darwinfoundation.org/en/checklist/

colnames(chartdf2) = c('Species','Galapagos..binary',"Islands..multistate","Carapace.length.cm." ,"Reference" ,"Galápagos..binary." ,"Islands..multistate..y", "Latitude","Longitude", "Island.Size..km.2." ,"Island.location.")

chartdf2$Latitude = as.double(chartdf2$Latitude)
chartdf2$Longitude = as.double(chartdf2$Longitude)
str(chartdf2)
justg = chartdf2[chartdf2$Galapagos..binary == 1,]

Ecuador <- gisco_get_countries(country = "ECU", resolution = 1)
g_plot = ggplot() +
  geom_sf(data = Ecuador, fill = "grey", alpha = 0.3) +
  geom_point(data = justg, 
             aes(x = Longitude, y = Latitude, size = Carapace.length.cm., color = Carapace.length.cm.),
             alpha = 0.7) + # alpha: opacity 
  scale_size_continuous(range = c(1, 12),name = 'Carapace Length (cm)') + # size legend
  scale_color_viridis_c(trans = "log", name = 'Carapace Length (cm)') + # color legend
  theme(legend.position = "right")+ # add legend
  labs(y= "Latitude", x = "Longitude", title ='Body Length by Island')+
  coord_sf(xlim = c(-92, -89), ylim = c(-2, 1)) # set window size

# set hover text
justg$hover_text <- paste("Species: ",justg$Species, "<br>Carapace Length: ", justg$Carapace.length.cm.)
# set plotly parameters:
plotly_galapagos <- ggplotly(g_plot) %>%
  style(
    hoverinfo = "text",  
    text = justg$hover_text )

plotly_galapagos

g_plot

library(htmlwidgets)
saveWidget(plotly_galapagos, "i_map_galapagos.html")
```


# correlation: island X bodysize
```{r}
chartdf_g = chartdf2[chartdf2$Island.Size..km.2. >0,]
cor(chartdf_g$Island.Size..km.2.,chartdf_g$Carapace.length)
```
# Linear Regression:
body size vs island size
```{r}
model = lm(Carapace.length.cm.~Island.Size..km.2., data = chartdf_g)
ggplot() +
  geom_point(data = chartdf_g, aes(x = Island.Size..km.2., y = Carapace.length.cm. )) +
  geom_abline(slope = coef(model)[["Island.Size..km.2."]], 
              intercept = coef(model)[["(Intercept)"]], color = 'darkgreen')+
  labs(y= 'Body Length', x = 'Island Size  (Km sq.)', title ='Body Length vs Island Size' )
```
# Model Summary
```{r}
summary(model)
```
# HiSSE
```{r}
suppressWarnings(library(hisse))
phy <- read.nexus("Supplementary File 3.tre")

five.rate.recon <- marginreconmisse(phy=phy, f=1, hidden.states=5,
pars=five.rate$solution, n.cores=1, aic=five.rate$aic)
```

```{r}
plot.misse.states(five.rate.recon, rate.param="net.div", show.tip.label=TRUE, type="phylogram", fsize=.25, legend="none",add = TRUE)
```

```{r}
tree <- read.nexus("Supplementary File 3.tre")
potential.combos <- generateMiSSEGreedyCombinations(max.param=4, vary.both=TRUE)
# running on one core slow but multiple would not work 
model.set <- MiSSEGreedy(tree, possible.combos=potential.combos, n.cores=1)

model.recons <- as.list(1:length(model.set))

for (model_index in 1:length(model.set)) {
nturnover <- length(unique(model.set[[model_index]]$turnover))
neps <- length(unique(model.set[[model_index]]$eps))
misse_recon <- MarginReconMiSSE(phy = model.set[[model_index]]$phy, f = 1,
hidden.states = nturnover,
pars = model.set[[model_index]]$solution,
AIC = model.set[[model_index]]$AIC)
model.recons[[model_index]] <- misse_recon
}

tip.rates <- GetModelAveRates(model.recons, type = c("tips"))
```


```{r}
write.csv(tip.rates,"tip.rates.csv")
(myTree3$tip.label)
# diversification = speciation - extinction rate
# extinction rate the same so truly a difference is speciation
colnames(tip.rates) = c('Species','state','turnover','net.div','speciation','extinct.frac','extinction')
tip.rates
```

```{r}
#Merge data together
chartdf.with.rates <- merge(chartdf2, tip.rates, by="Species")
# write out for Hubert
chartdf.with.rates
write.csv(chartdf.with.rates,"tip.rates.combo.csv")
```
```{r}
# average rates for Galapagos
mean(tip.rates[20:30,]$net.div)
# average rates for not Galapagos 
mean(tip.rates[-(20:30),]$net.div)
```



